### 2019-08-28 Writeup/Paper/sections/introduction.tex
Changes; - Implemted changes from Sia - Added abstract and acknowledgements

\chapter{Introduction}
\pagenumbering{arabic}
\label{cha:Introduction}

The increasing use of public-key cryptography by instant messaging and secure email means ensuring confidentiality is an ever more important task.

One of the most significant risks to the security of the communication channel is a Man-in-the-middle (MiTM) attack. A MiTM attack involves an attacker impersonating one or both sides of a connection. MiTM attacks can entirely circumvent the encryption as it allows an attacker to read all of the encrypted data.

A countermeasure for the threat of MiTM attacks is the verification of each parties’ fingerprint. A fingerprint is a small string of characters that is unique to each key and, thus, can be used to identify.

Fingerprints can come in several different encodings such as Hexadecimal, words, and even procedurally generated avatars. Previous research has shown that the average human can only hold around 7-digits ($\pm 2$) worth of data in their working memory\cite{miller1956magical}. Consequently, this makes the designing of user-friendly schemes a task of utmost importance.

Humans are commonly considered the most vulnerable part of any computer system. Solutions, therefore, have been proposed to remove this manual verification with examples such as PGP's web-of-trust\cite{callas1998openpgp} and Namecoin\cite{kalodner2015empirical}. However, these suffer from user adoption due to perceived complexity. Manual verification is, consequently, left in a difficult position, due in part to proposed solutions needing to sacrifice either usability or security. Therefore, research into improving or creating a more secure and user-friendly fingerprint encoding remains an important task.

One proposed user-friendly scheme is Pretty Easy Privacy's (p$\equiv$p) "Trustwords". Trustwords is an implementation of word fingerprint mapping with a emphasis on usability. This increase in usability is achieved by the user comparing a reduced number of words. This usability boost, however, comes at a cost of a much larger word list.

This report evaluates the security and feasibility of attacking Trustwords. The motivation for this is due in part to the shortage of justification behind the size and features of the chosen wordlist.

\newpage
The report is structured as follows:

\begin{itemize}
    \item \textbf{Chapter 2} provides any required background knowledge alongside a review of all relevant literature.

    \item \textbf{Chapter 3} formally defines the aim of this report alongside the research questions the project aims to answer.

    \item \textbf{Chapter 4} documents the proposed designs required to answer the research questions proposed in Chapter 3.

    \item \textbf{Chapter 5} discusses the significant implementation details of the designs formalized in Chapter 4.

    \item \textbf{Chapter 6} presents and analyses all experimentation results of the project.

    \item \textbf{Chapter 7} evaluates and concludes the success of the project alongside a discussion into proposed further work. The research aims proposed earlier in Chapter 3 are used as evaluative metrics.
\end{itemize}

>>>>

\section{Introduction}

The increasing use of public-key cryptography by instant messaging and secure email means ensuring confidentiality is an ever more important task.
One of the most significant risks to the security of the communication channel is a Man-in-the-middle (MiTM) attack. A MiTM attack involves an attacker impersonating one or both sides of a connection. MiTM attacks can entirely circumvent the encryption as it allows an attacker to read all of the encrypted data.
A countermeasure for the threat of MiTM attacks is the verification of each parties’ fingerprint. A fingerprint is a small string of characters that is unique to each key and, thus, can be used to identify.
Fingerprints can come in several different encodings such as Hexadecimal, words, and even procedurally generated avatars. Previous research has shown that the average human can only hold around 7-digits ($\pm 2$) worth of data in their working memory\cite{miller1956magical}. Consequently, this makes the designing of user-friendly schemes a task of utmost importance.
Humans are commonly considered the most vulnerable part of any computer system. Solutions, therefore, have been proposed to remove this manual verification with examples such as PGP's web-of-trust\cite{callas1998openpgp} and Namecoin\cite{kalodner2015empirical}. However, these suffer from user adoption due to perceived complexity. Manual verification is, consequently, left in a difficult position, due in part to proposed solutions needing to sacrifice either usability or security. Therefore, research into improving or creating a more secure and user-friendly fingerprint encoding remains an important task.
One proposed user-friendly scheme is Pretty Easy Privacy's (p$\equiv$p) "Trustwords". Trustwords is an implementation of word fingerprint mapping with a emphasis on usability. This increase in usability is achieved by the user comparing a reduced number of words. This usability boost, however, comes at a cost of a much larger word list.
This report evaluates the security and feasibility of attacking Trustwords. The motivation for this is due in part to the shortage of justification behind the size and features of the chosen wordlist.

### 2019-08-28 Writeup/Thesis/chapters/design.tex
Changes; - Updated a global code style for the paper - Fixed some more supervisor issues - Small changes
Changes; - Implemted changes from Sia - Added abstract and acknowledgements

The attack on Trustwords involves generating "near-collision" keys. 
This attempts to provide the base to answer Research Questions \ref{goal:numberOfTrustwords}, \ref{goal:complexity} and \ref{goal:hardwareRequired}.

@@

\section{Design of GreenOnion}
\label{sec:greenDesign}
In order to answer Research Questions \ref{goal:complexity} and \ref{goal:hardwareRequired} fully, it is necessary to implement an actual tool to generate actual keys.

@@

\subsection{Metric performance}
\label{exp:metric}
This section answers Research Question \ref{goal:bestMetrics}. To reduce the number of metrics to assess in the later rounds, the performance of similarity metrics required comparison. The thinning out of metrics is required due to limited resources. Therefore, possible further work could involve repeating this work with a much more varied selection of metrics. As explained in a previous section (See Section \ref{sec:metrics}) the chosen metrics up for assessment are Soundex, Metaphone, NYSIIS, Levenshtein and Phonetic vectors.

@@

\subsubsection{Comparisons}
Each comparison contains a match-pair. These matches are generated by running the similarity metric over all the pairings in the Trustword dictionary, if the pair meets the criteria discussed in the previous section it is determined a match. Potential matches are then sampled from their respective lists.

@@

@@

\todo{Include explination into how audio is generated}

>>>>

The attack on Trustwords involves generating "near-collision" keys. 
This attempts to provide the base to answer Research Questions \ref{goal:numberOfTrustwords} and \ref{goal:complexity}.

@@

\section{Design of GreenOnion}
\label{sec:greenDesign}
In order to answer Research Questions \ref{goal:complexity} fully, it is necessary to implement an actual tool to generate actual keys.

@@

\subsection{Metric performance}
\label{exp:metric}
This section answers Research Question \ref{goal:phoneticSimilarity}. To reduce the number of metrics to assess in the later rounds, the performance of similarity metrics required comparison. The thinning out of metrics is required due to limited resources. Therefore, possible further work could involve repeating this work with a much more varied selection of metrics. As explained in a previous section (See Section \ref{sec:metrics}) the chosen metrics up for assessment are Soundex, Metaphone, NYSIIS, Levenshtein and Phonetic vectors.

@@

\subsubsection{Comparisons}
\label{sec:exp1_comparison}
Each comparison contains a match-pair. These matches are generated by running the similarity metric over all the pairings in the Trustword dictionary, if the pair meets the criteria discussed in the previous section it is determined a match. Potential matches are then sampled from their respective lists.

@@

\newpage

@@

### 2019-08-28 Writeup/Thesis/chapters/evaluation.tex
Changes; - Implemted changes from Sia - Added abstract and acknowledgements

Research Question \ref{goal:phoneticSimilarity} asks: \textit{``What are the different ways phonetic similarity can be quantified?''}. This paper has discussed several metrics used to quantify phonetic similarity (See Section \ref{sec:metrics}). Most metrics were chosen due to their high utilization and occurrence in literature. The only exception to this is the presence of Phonetic Vectors that were chosen due to their unique qualities. This project has succeeded in presenting a comparison of ways to compare similarity metrics. A more in-depth comparison of all possible metrics is a substantial task and would require a project with this sole dedication.

\section{Research Question \ref{goal:bestMetrics}}
Research Question \ref{goal:bestMetrics} asks: \textit{``Out of a chosen set of metrics, which are the most effective?''}. 
The paper has presented the performance of a suite of algorithms that are mostly present in alternative literature. Performance of these metrics was quantified in a direct study that selected the top three from a set of five. These newly selected metrics were then assessed in an attack context with their performance being indirectly quantified. From the historical set of metrics, Levenshtein was the best performer with participants rating its matches as substantially better. Levenshtein was also the best performer in the attack context with it having the best average attack success rate. This performance, therefore, seems to corroborate work showing that words with similar spelling tend to share phonetic similarity \cite{hettiarachchi2012sparcl}. This area, however, requires further work to provide more conclusive results.

\section{Research Question \ref{goal:complexity}}
Research Question \ref{goal:complexity}: \textit{``What are the time and computation complexity required to generate a ’similar’ keys for a targeted key pair?''} has been discussed by the project, the actual generation of keys as shown in Section \ref{sec:key_gen} where actual keys were generated. Furthermore, the distribution of key permutations on a set of real-world keys were discussed in Section \ref{sec:averagePerms} and \ref{sec:vulnKeys}. To achieve this, the tool GreenOnion was developed where inspiration was gathered from a currently available tool known as Scallion. Experimentation results were gathered and displayed in Section \ref{sec:SvG} that showed the significant increase in the number of near-collision keys GreenOnion can test for with GreenOnion being able to handle in excess of 1.5 million keys. This capability is in comparison to the maximum of 40 keys Scallion can compare.

\section{Research Question \ref{goal:hardwareRequired}}
Research Question \ref{goal:hardwareRequired} states: \textit{``What kind of hardware is required to compute a matching key?''}. Through the computation of actual keys presented in Section \ref{sec:key_gen} and the theoretical calculation into key computation in some cases, it has been shown that a single mid-range GPU\footnote{AMD RX-480} can be used to compute actual keys.

@@

Furthermore, comparison of these results to similar literature shows a considerable increase in success rates for the attack present in this paper. All alternative literature used a small dictionary of around 1024 words. Therefore, this is highly suggestive that the issue lies with the design of Trustwords. Therefore, we believe that 4 Trustwords as a minimum has not provided a basic level of security towards users.

>>>>

Research Question 1 asks: \textit{“What are the best performing schemes to quantify phonetic similarity?”}. This paper has discussed several metrics used to quantify phonetic similarity (See Section \ref{sec:metrics}). Most metrics were chosen due to their high utilization and occurrence in literature. The only exception to this is the presence of Phonetic Vectors that were chosen due to their unique qualities. Performance of these metrics were then quantified in a direct study that selected the top three from a set of five (See Section \ref{exp:metric}). These newly selected metrics were then assessed in an attack context with their performance being indirectly quantified (See Section \ref{sec:exp2_design}).

The top three metrics (in descending order) were: Levenshtein, NYSIIS and Metaphone. We found that Levenshtein was the most aligned with how people rate phonetically similarity. Levenshtein had a considerable proportion of 4 and 5 ratings (64\%) compared to the alternatives. Levenshtein was also the best performer in the attack context with it having the best attack success rate in all attacks. This corroborates work showing that words with similar spelling tend to share phonetic similarity \cite{hettiarachchi2012sparcl}.
Alongside this, we found Soundex was the least aligned with the human model of similarity with an average rating of just 2 out of 5. This corroborates findings that highlighted major issues with the accuracy of Soundex’s phonetic representation (See Section \ref{sec:soundex}).

We believe the outcome of the first experiment were inflated by the way the results were presented (as discussed in Section \ref{sec:exp1_considerations}). However, as the performance of the metrics in the second experiment emulated the first with Levenshtein being the best performer alongside Metaphone being the worst, shows that the results of the first study maintains a desirable level of accuracy. Improvements can be made with the proposed design discussed in Section \ref{sec:exp1_considerations}, but, we believe the first experiment has achieved its goal of providing a low cost way to reduce the number of selected metrics for the subsequent study.

\section{Research Question \ref{goal:complexity}}
Research Question \ref{goal:complexity}: \textit{``What are the time and computation complexity required to generate a ’similar’ keys for a targeted key pair?''} has been discussed by the project, the actual generation of keys as shown in Section \ref{sec:key_gen} where actual keys were generated. Furthermore, the distribution of key permutations on a set of real-world keys were discussed in Section \ref{sec:averagePerms} and \ref{sec:vulnKeys}. To achieve this, the tool GreenOnion was developed where inspiration was gathered from a currently available tool known as Scallion. Experimentation results were gathered and displayed in Section \ref{sec:SvG} that showed the significant increase in the number of near-collision keys GreenOnion can test for with GreenOnion being able to handle in excess of 1.5 million keys. This capability is in comparison to the maximum of 40 keys Scallion can compare. We believe this is due to the implementation of the bloom filter. Its unique characteristics of $O(k)$ complexity regardless of number of elements in the structure results in a huge number of concurrent checks to occur.

@@

Furthermore, comparison of these results to similar literature shows a considerable increase in success rates for the attack present in this paper. We believe this is due to the consideration of `similarity' when forming attacks. This was an aspect missing from the literature.
The effect of `similarity' is further enhanced due to the  weaknesses present in Trustwords such as homophones, dual mappings and the wordlist size. Therefore, we believe that 4 Trustwords as a minimum has not provided a basic level of security towards users.

### 2019-08-28 Writeup/Thesis/chapters/experiments.tex
Changes; - Implemted changes from Sia - Added abstract and acknowledgements

@@

Figure \ref{tab:scallion_speed} shows the speed decline as the number of potential keys checked in one go increases for Scallion. Any number of keys over 40 resulted in a crash for Scallion. GreenOnion's speed is almost perfectly consistent. It is slower up until  around 5 keys. This consistency is due to the overhead of the bloom filter; however, as the number of keys increase, the bloom filter's benefits become apparent. In this test, the bloom filter size was kept consistent at exactly 5,000 elements.

@@

Work by \textbf{R. Kainda, I. Flechais} and \textbf{A. Roscoe, \textit{et al.}}\cite{kainda2009usability} (previously discussed in Chapter \ref{cha:LiteratureReview}) is the most comparable to this experiment. They were the only study to use exactly 4 words but differed on the size of the dictionary (1024 words) and how a near-match were calculated. Near-matches were a difference in a single word, but without the consideration for similarity, this is the unique aspect considered with this experiment.

>>>>

\todo{Talk about the exception}

@@

Figure \ref{tab:scallion_speed} shows the speed decline as the number of potential keys checked in one go increases for Scallion. Any number of keys over 40 resulted in a crash for Scallion. This crash is solely due to an  implementation issue.

\todo{Fix implementation issue}

GreenOnion's speed is almost perfectly consistent. It is slower up until  around 5 keys. This consistency is due to the overhead of the bloom filter; however, as the number of keys increase, the bloom filter's benefits become apparent. In this test, the bloom filter size was kept consistent at exactly 5,000 elements.

@@

Work by \textbf{R. Kainda, I. Flechais} and \textbf{A. Roscoe}\cite{kainda2009usability} (previously discussed in Chapter \ref{cha:LiteratureReview}) is the most comparable to this experiment. They were the only study to use exactly 4 words but differed on the size of the dictionary (1024 words) and how a near-match were calculated. Near-matches were a difference in a single word, but without the consideration for similarity, this is the unique aspect considered with this experiment.
